flowCore: a Bioconductor package for high throughput flow cytometry

### 2013-02-06 inst/doc/HowTo-flowCore.Rnw
Fixed typos in the vignette.
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/flowCore@43485 bc3139a8-67e5-0310-9ffc-ced21a209358
initial checkin

\noindent \textbf{Results} We hope that those proposed data structures
will the base for the development of many tools for the analysis of
high throughput flow cytometry.\\

@@

You can extract a \Rclass{flowFrame} from a \Rclass{flowSet} object in
the usual way using the \Rfunction{[[} or \Rfunction{\$} extraction
  operators. On the other hand using he \Rfunction{[} extraction
  operator returns a new \Rclass{flowSet} by \textbf{copying} the
  environment. However, simply assigning the \Rclass{flowFrame} to a
  new variable will \textbf{not} copy the contained frames.

@@

<<fsApply1, echo=true, results=verbatim>>=
fsApply(fs, each_col, median)
@

which is equivalent to the less readable

<<fsApply2, echo=true,results=verbatim>>=
fsApply(fs,function(x) apply(x, 2, median), use.exprs=TRUE)
@

In this case, the \Robject{use.exprs} argument is not required in the
first case because \Rfunction{each\_col} and \Rfunction{each\_row} are
methods and have been defined to work on \Rclass{flowFrame} objects by
first extracting the intensity data.

>>>>

\noindent \textbf{Results} We hope that those proposed data structures
will be the base for the development of many tools for the analysis of
high throughput flow cytometry.\\

@@

You can extract a \Rclass{flowFrame} from a \Rclass{flowSet} object in
the usual way using the \Rfunction{[[} or \Rfunction{\$} extraction
  operators. On the other hand using the \Rfunction{[} extraction
  operator returns a new \Rclass{flowSet} by \textbf{copying} the
  environment. However, simply assigning the \Rclass{flowFrame} to a
  new variable will \textbf{not} copy the contained frames.

@@

<<fsApply1, echo=true, results=verbatim>>=
fsApply(fs, each_col, median)
@
%
which is equivalent to the less readable

<<fsApply2, echo=true,results=verbatim>>=
fsApply(fs,function(x) apply(x, 2, median), use.exprs=TRUE)
@
%
In this case, the \Robject{use.exprs} argument is not required in the
first case because \Rfunction{each\_col} and \Rfunction{each\_row} are
methods and have been defined to work on \Rclass{flowFrame} objects by
first extracting the intensity data.

### 2013-02-06 paper/bmc_article.tex
initial checkin

\subsubsection*{Transformation and compensation}
Data transformation is essential for both data visualization and
modelling \cite{lo2008agf}. The major transformations that are
routinely used in FCM analysis have been implemented in
\Rpackage{flowCore} (\textit{e.g.,} log, bi-exponential or arcsinh,
see Table~\ref{table1} for a complete list). Furthermore, the design
of the R language makes it easy to define arbitrary functions to apply
to the data of individual \Rclass{flowFrames} or entire
\Rclass{flowSets}, respectively. Compensation, that corrects for
fluorescence spillover originating from the inherent overlap of
emission spectra from antibody fluorescent labels, is available for
both \Rclass{flowFrames} and \Rclass{flowSets}. In addition, the
software offers functionality to compute spillover or compensation
matrices from a set of appropriate single stain controls.

@@

The ability to programmatically access gates is a prerequisite for
semi-automated or automated gating. By utilizing an unified interface
for all different types of gates, the user is able to subset data sets
as well as to create summary statistics, as for instance the
proportions of events falling in a single gate or in a combination of
gates. Complex combinations and hierarchies of gates can be captured
in multi-step gating strategies. The definition of gates in
\Rpackage{flowCore} follows the Gating-ML CR, thus any
\Rpackage{flowCore} gating strategy can be reproduced by any other
software that also adheres to the standard and \textit{vice versa}.

@@

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
  \subsection*{Table 1 - Data transformations}
    \begin{table}[ht]
      \caption{\label{table1} Data transformations implemented in \Rpackage{flowCore}. Within these formulas, $x$ is the variable
        corresponding to value being transformed, $y$ is the result of the
        transformation, $a$, $b$, $c$, $d$, $f$, $p$, $m$, $T$, and $w$, are
        constants affecting the transformation function, $e$ is the base of
        the natural logarithm, and $root$ is a root-finding algorithm. Other transformations can easily be implemented in R.}
      \begin{center}
        \begin{tabular}{|l|l|}
          \hline
          \multicolumn{2}{|c|}{Data Transformations} \\
          \hline
          linear & $ax + b$ \\
          quadratic & $ax^2 + bx + c$ \\
          natural logarithm & $log_e(x)(r/d)$ \\
          logarithm & $log_b(x)(r/d)$ \\
          biexponential & $ae^{(b*x)}-ce^{(-d+x)}+f$ \\
          logicle& $Te^{-(m-w)}(e^{(x-w)}-p^2e^{-(x-w)/p}+p^2-1)$ \\
          truncate & $x_{x{\leq}a} = a$ \\
          scale & $(x-a)/(b-a)$ \\
          arcsinh & $arcsinh(a + bx)+c$ \\
          \hline
        \end{tabular}
      \end{center}
    \end{table}

>>>>

\subsubsection*{Transformation and compensation}
Data transformation is essential for both data visualization and
modelling \cite{lo2008agf}. The major transformations that are
routinely used in FCM analysis have been implemented in
\Rpackage{flowCore} (\textit{e.g.,} log, bi-exponential, arcsinh or
logicle \cite{Parks:2006}, see Table~\ref{table1} for a complete
list). Furthermore, the design of the R language makes it easy to
define arbitrary functions to apply to the data of individual
\Rclass{flowFrames} or entire \Rclass{flowSets},
respectively. Compensation, that corrects for fluorescence spillover
originating from the inherent overlap of emission spectra from
antibody fluorescent labels, is available for both \Rclass{flowFrames}
and \Rclass{flowSets}. In addition, the software offers functionality
to compute spillover or compensation matrices from a set of
appropriate single stain controls.

@@

The ability to programmatically access gates is a prerequisite for
semi-automated or automated gating. By utilizing an unified interface
for all different types of gates, the user is able to subset data sets
as well as to create summary statistics, as for instance the
proportions of events falling in a single gate or in a combination of
gates. Complex combinations and hierarchies of gates can be captured
in multi-step gating strategies. The definition of gates in
\Rpackage{flowCore} follows the Gating Markup Language Candidate
Recommendation \cite{SpidlenInPressCytometryA}, thus any
\Rpackage{flowCore} gating strategy can be reproduced by any other
software that also adheres to the standard and \textit{vice versa}.

@@

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
  \subsection*{Table 1 - Data transformations}
    \begin{table}[ht]
      \caption{\label{table1} Data transformations implemented in \Rpackage{flowCore}. Within these formulas, $x$ is the variable
        corresponding to value being transformed, $a$, $b$, $c$, $d$, $f$, $p$, $m$, $T$, and $w$, are
        constants affecting the transformation function, $e$ is the base of the natural logarithm (see \cite{Parks:2006} for details on the logicle transformation). Other transformations can easily be implemented in R.}
      \begin{center}
        \begin{tabular}{|l|l|}
          \hline
          \multicolumn{2}{|c|}{Data Transformations} \\
          \hline
          linear & $ax + b$ \\
          quadratic & $ax^2 + bx + c$ \\
          natural logarithm & $log_e(x)(r/d)$ \\
          logarithm & $log_b(x)(r/d)$ \\
          biexponential & $ae^{(b*x)}-ce^{(-d+x)}+f$ \\
          logicle& $Te^{-(m-w)}(e^{(x-w)}-p^2e^{-(x-w)/p}+p^2-1)$ \\
          truncate & $x_{x{\leq}a} = a$ \\
          scale & $(x-a)/(b-a)$ \\
          arcsinh & $arcsinh(a + bx)+c$ \\
          \hline
        \end{tabular}
      \end{center}
    \end{table}
